\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis

### 2021-08-02 manuscript.Rnw
Minor edits to prose; add references for R packages

The functionality of existing packages can generally be separated into three categories: estimation of model parameters, simulation of model outcomes, and decision analysis. Nearly all packages mentioned above can be used to perform decision analysis in some capacity: \pkg{BCEA}, \pkg{SAVI}, \pkg{dampack}, \pkg{EVSI}, and \pkg{heemod} all contain features for post-processing the output of simulation models in order to quantify decision uncertainty and conduct value of information analyses. On the other hand, there is, to our knowledge, only one package, \pkg{heemod}, that is designed for simulation of health economic models. \pkg{heemod} can be used to develop Markov cohort models and can incorporate parameter uncertainty for a PSA. Finally, it is important to note that one of, it not the primary advantages of using \proglang{R} for cost-effectiveness modeling is the rich set of packages available for statistical analysis. While any number of these can be used for model parameter estimation---as will be partially demonstrated in this paper---\pkg{survHE} is unique in that it was designed specifically for use in health economic evaluations and consequently contains features such as convenient comparisons of multiple parametric survival models and simulation of model parameters for PSA.

@@

Of the existing packages, \pkg{hesim} is most similar to \pkg{heemod} in that it is designed primarily for simulation of economic models, but can also be used to conduct a CEA. That said, there are some notable differences. First, \pkg{hesim} can be used for individual-level simulations, whereas \pkg{heemod} can only simulate cohort-level models. Second, all outcomes in \pkg{hesim} are simulated across parameter samples for a PSA, treatment strategies, and patient subgroups in a vectorized fashion. This not only increases speed because outcomes are simulated at the \proglang{C++} level, but also makes it easier to account for patient heterogeneity and compare a large number of treatment strategies. Third, since all model types in \pkg{hesim} can be parameterized by fitting statistical models in \proglang{R}, it is easier to integrate the statistical analysis and economic model structure.

@@

The three types of economic models described in Section~\ref{sec:model-taxonomy} are implemented as the \pkg{R6} classes \code{CohortDtstm}, \code{IndivCtstm}, and \code{Psm}. Each class contains methods for simulating disease progression, QALYs, and costs. These methods are made possible by separate \pkg{R6} classes for the disease, utility, and cost submodels. The remainder of this section describes the framework for parameterization and simulation. The cost-effectiveness framework is described in Section~\ref{sec:cea}.

@@

The parameters of a survival model are stored in a \code{params_surv} object and a \code{params_surv_list} can be used to store the parameters of multiple survival models. The latter is useful for storing the parameters of a multi-state model or the independent survival models required for a PSM. The parameters of a multinomial logistic regression are stored in a \code{params_mlogit} object and can be created by fitting a model for each row in a transition probability matrix with \code{nnet::multinom()}.

\code{tparams_transprobs} objects store transition probability matrices that have been predicted for each PSA sample, treatment strategy, patient from the target population, and optionally time interval (to allow for time inhomogeneous Markov models). Transition probabilities can be predicted from a fitted multi-state model using the \pkg{msm} package or constructed ``by hand'' in a custom manner.

@@

For a given subgroup, QALYs and costs are simulated for each PSA sample, treatment strategy, and health state. They are summarized with the \code{$summarize()} method, which first sums across health states and then computes a (weighted) average across patients, by PSA sample, treatment strategy, and subgroup. The output is a generic cost-effectiveness object (of class \code{ce}) that can be used to perform CEA as demonstrated in Section~\ref{sec:cea}. Here, we use the \code{summary()} method to summarize the results and make use of the ``pipes'' from the \pkg{magrittr} to nicely format the output.

@@

A third helpful feature that we did not mention is that a PSM can be extended to $N$ health states. The 3-state oncology case might, for instance, be extended to account for adverse events or multiple lines of therapy. One related application is the quality-adjusted time without symptoms or toxicity (Q-TWiST) method \citep{goldhirsch1989costs, lenderking1994eval}, which partitions patients into time spent (i) having toxic side effects, (ii) without symptoms or toxicities, and (iii) after progression. An example four-state model is provided on the package website.

>>>>

The functionality of existing packages can generally be separated into three categories: estimation of model parameters, simulation of model outcomes, and decision analysis. Nearly all packages mentioned above can be used to perform decision analysis in some capacity: \pkg{BCEA}, \pkg{SAVI}, \pkg{dampack}, \pkg{EVSI}, and \pkg{heemod} all contain features for post-processing the output of simulation models in order to quantify decision uncertainty and conduct value of information analyses. On the other hand, there is, to our knowledge, only one package, \pkg{heemod}, that is designed for simulation of health economic models. \pkg{heemod} can be used to develop Markov cohort models and can incorporate parameter uncertainty for a PSA. Finally, it is important to note that one of, it not the primary advantage of using \proglang{R} for cost-effectiveness modeling is the rich set of packages available for statistical analysis. While any number of these can be used for parameter estimation, \pkg{survHE} is unique in that it was designed specifically for use in health economic evaluations and consequently contains specialized features such as convenient comparisons of multiple parametric survival models and simulation of parameters for PSA.

@@

Of the existing packages, \pkg{heemod} is most similar to \pkg{hesim} in that it is designed primarily for simulation of economic models, but can also be used to conduct a CEA. That said, there are some notable differences. First, \pkg{hesim} can be used for individual-level simulations, whereas \pkg{heemod} can only simulate cohort-level models. Second, all outcomes in \pkg{hesim} are simulated across parameter samples for a PSA, treatment strategies, and patients in a vectorized fashion. This not only increases speed because outcomes are simulated at the \proglang{C++} level, but also makes it easier to account for patient heterogeneity and compare a large number of treatment strategies. Third, since all model types in \pkg{hesim} can be parameterized by fitting statistical models in \proglang{R}, it is easier to integrate the statistical analysis and economic model structure.

@@

The three types of economic models described in Section~\ref{sec:model-taxonomy} are implemented as the \pkg{R6} \citep{chang2020} classes \code{CohortDtstm}, \code{IndivCtstm}, and \code{Psm}. Each class contains methods for simulating disease progression, QALYs, and costs. These methods are made possible by separate \pkg{R6} classes for the disease, utility, and cost submodels. The remainder of this section describes the framework for parameterization and simulation. The cost-effectiveness framework is described in Section~\ref{sec:cea}.

@@

The parameters of a survival model are stored in a \code{params_surv} object and a \code{params_surv_list} can be used to store the parameters of multiple survival models. The latter is useful for storing the parameters of a multi-state model or the independent survival models required for a PSM. The parameters of a multinomial logistic regression are stored in a \code{params_mlogit} object and can be created by fitting a model for each row in a transition probability matrix with \code{nnet::multinom()} \citep{venables2002}.

\code{tparams_transprobs} objects store transition probability matrices that have been predicted for each PSA sample, treatment strategy, patient from the target population, and optionally time interval (to allow for time inhomogeneous Markov models). Transition probabilities can be predicted from a fitted multi-state model using the \pkg{msm} package \citep{jackson2011} or constructed ``by hand'' in a custom manner.

@@

For a given subgroup, QALYs and costs are simulated for each PSA sample, treatment strategy, and health state. They are summarized with the \code{$summarize()} method, which first sums across health states and then computes a (weighted) average across patients, by PSA sample, treatment strategy, and subgroup. The output is a generic cost-effectiveness object (of class \code{ce}) that can be used to perform CEA as demonstrated in Section~\ref{sec:cea}. Here, we use the \code{summary()} method to summarize the results and make use of the ``pipes'' from the \pkg{magrittr} \citep{bache2020} package to nicely format the output.

@@

A third feature that we did not mention is that a PSM can be extended to $N$ health states. The 3-state oncology case might, for instance, be extended to account for adverse events or multiple lines of therapy. One related application is the quality-adjusted time without symptoms or toxicity (Q-TWiST) method \citep{goldhirsch1989costs, lenderking1994eval}, which partitions patients into time spent (i) having toxic side effects, (ii) without symptoms or toxicities, and (iii) after progression. An example four-state model is provided on the package website.
